Pitfall of Prediction-orienterd Machine Learning Methods and Economic Value of Statistical Efficiency in On-farm Precision Experimentation

### 2022-02-25 Writing/manuscript.rmd
Updates: (1) best-performing model count table, (2) EONR ratio kernel, (3) pairwise t-test tables

# /*----------------------------------*/
#' ##  data preparation
# /*----------------------------------*/
#' load results data
est_data <-
  readRDS(here("Shared", "Results", "est_result_ls.rds")) %>%
  rbindlist() %>%
  dplyr::select(model, perform, field_col) %>%
  unnest(perform) %>%
  data.table()
#' rename columns
est_data <- est_data %>%
  .[, .(field_col, model, sim, profit, rmse_train, rmse_cv, rmse_eonr)] %>%
  #---field size---
  .[, field_size := paste0(round(field_col * 72 * 6^2 / 10000, 1), " ha")] %>%
  .[, field_size := factor(field_size, levels = c("9.3 ha", "18.7 ha", "37.3 ha"))] %>%
  #---model name---
  .[model == "brf", model := "BRF"] %>%
  .[model == "rf", model := "RF"] %>%
  .[model == "rf_perfect", model := "RF_b"] %>%
  .[model == "lm", model := "OLS"] %>%
  .[model == "ser_50", model := "SER"] %>%
  .[, model := factor(model, levels = c("RF", "RF_b", "BRF", "OLS", "SER"))]
#' mean and sd
mean_data <- est_data %>%
  .[, .(
    profit = mean(profit) %>% round(2),
    profit_sd = sd(profit) %>% round(2),
    rmse_yield = mean(rmse_cv) %>% round(2),
    rmse_yield_sd = sd(rmse_cv) %>% round(2),
    rmse_eonr = mean(rmse_eonr) %>% round(2),
    rmse_eonr_sd = sd(rmse_eonr) %>% round(2)
  ),
  by = c("field_size", "model")
  ] %>%
  .[order(field_size, model), ]
#' load illustration data (single simulation)
illus_data <-
  readRDS(here("Shared", "Results", "illus_data.rds"))

@@

```{r yield-curves, fig.cap="Estimated yield response curves at a specific location", fig.height=8.5, fig.width=7, echo=FALSE, cache=FALSE}
include_graphics(here("Shared", "Graphs", "802_yield.png"))
```

@@

@@

```{r extreme-percent, echo=FALSE}
# knitr::kable(mean_data, format = "markdown", caption = "Summary Statistics")
```

>>>>

# /*----------------------------------*/
#' ##  data preparation
# /*----------------------------------*/
#' load results data
est_data <-
  readRDS(here("Shared", "Results", "est_result_ls.rds")) %>%
  rbindlist() %>%
  dplyr::select(model, perform, field_col) %>%
  unnest(perform) %>%
  data.table() %>% 
  .[, .(field_col, model, sim, profit, rmse_train, rmse_cv, rmse_eonr)] %>%
  #---field size---
  .[, field_size := paste0(round(field_col * 72 * 6^2 / 10000, 1), " ha")] %>%
  .[, field_size := factor(field_size, levels = c("37.3 ha", "18.7 ha", "9.3 ha"))] %>%
  #---model name---
  .[model == "brf", model := "BRF"] %>%
  .[model == "rf", model := "RF"] %>%
  .[model == "rf_perfect", model := "RF_b"] %>%
  .[model == "lm", model := "OLS"] %>%
  .[model == "ser_50", model := "SER"] %>%
  .[, model := factor(model, levels = c("RF", "RF_b", "BRF", "OLS", "SER"))]

#' mean and sd
mean_data <- est_data %>%
  .[, .(
    profit = mean(profit) %>% round(2),
    profit_sd = sd(profit) %>% round(2),
    rmse_yield = mean(rmse_cv) %>% round(2),
    rmse_yield_sd = sd(rmse_cv) %>% round(2),
    rmse_eonr = mean(rmse_eonr) %>% round(2),
    rmse_eonr_sd = sd(rmse_eonr) %>% round(2)
  ),
  by = c("field_size", "model")
  ] %>%
  .[order(field_size, model), ]

#' aunit level data
aunit_data <-
    readRDS(here("Shared", "Results", "aunit_data.rds"))

@@

```{r yield-curves, fig.cap="Estimated yield response curves at a specific location", fig.height=5.5, fig.width=5, echo=FALSE, cache=FALSE}
include_graphics(here("Shared", "Graphs", "802_yield.png"))
```

@@

```{r eonr-ratio-kernel, fig.cap="Ratio of Estimated EONR to True EONR", fig.height=4.5, fig.width=5, echo=FALSE, cache=FALSE}
eonr_ratio_kernel
```

@@

```{r count-table, echo=FALSE, cache=FALSE}
prt_table <- est_data %>% 
    .[order(sim, model), ] %>% 
    .[, y := (rmse_cv==min(rmse_cv)) %>% as.numeric(), by = .(sim, field_size)] %>% 
    .[, eonr := (rmse_eonr==min(rmse_eonr)) %>% as.numeric(), by = .(sim, field_size)] %>% 
    .[, pi := (profit==max(profit)) %>% as.numeric(), by = .(sim, field_size)] %>% 
    .[, .(`yield` = (sum(y)/.N*100) %>% round(2),
          `EONR` = (sum(eonr)/.N*100) %>% round(2),
          `profit` = (sum(pi)/.N*100) %>% round(2)), by = .(field_size, model)]
# datasummary(
#         field_size * (yield + EONR + profit) ~ factor(model) * (Mean), 
#         data = count_table,
#         title = 'Percentage (%) of each model being the best-performing among the five models in a single simulation round',
#         notes = c('A note at the bottom of the table.')
#     )
count_table <- prt_table %>% 
    #=== Wide to Long: melt()
    melt(id.vars = c('field_size','model')) %>% 
    #=== Long to Wide: dcast()
    dcast(field_size + variable ~ model, value.var = "value")
knitr::kable(count_table, format = "markdown",
             caption = "Percentage (%) of each model being the best-performing among the five models in a single simulation round")
```

```{r pair-table-yield, echo=FALSE, cache=FALSE}
model_list <- unique(est_data$model) %>% .[order(.)] %>% as.character()
pair_df <- data.frame()
for(i in 1:length(model_list)){
    for(j in 1:length(model_list)){
        pi <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[i]) %>% 
            pull(rmse_cv)
        pj <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[j]) %>% 
            pull(rmse_cv)
        pv <- t.test(pi, 
                     pj, 
                     paired = TRUE)

        pair_df[model_list[i], model_list[j]] <- 
            round(pv$estimate, 2)
        pair_df[as.character(i), model_list[j]] <- 
            paste0("(", round(pv$p.value, 3), ")")
    }
}
pair_df[is.na(pair_df)] <- ""
knitr::kable(pair_df, format = "markdown",
             caption = "Pairwise t-test of predicted yield RMSE between models, Large field (37.3 ha), values are the difference in RMSE (row vs. index) (kg/ha), p-values in parenthese")
```

```{r pair-table-eonr, echo=FALSE, cache=FALSE}
model_list <- unique(est_data$model) %>% .[order(.)] %>% as.character()
pair_df <- data.frame()
for(i in 1:length(model_list)){
    for(j in 1:length(model_list)){
        pi <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[i]) %>% 
            pull(rmse_eonr)
        pj <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[j]) %>% 
            pull(rmse_eonr)
        pv <- t.test(pi, 
                     pj, 
                     paired = TRUE)

        pair_df[model_list[i], model_list[j]] <- 
            round(pv$estimate, 2)
        pair_df[as.character(i), model_list[j]] <- 
            paste0("(", round(pv$p.value, 3), ")")
    }
}
pair_df[is.na(pair_df)] <- ""
knitr::kable(pair_df, format = "markdown",
             caption = "Pairwise t-test of predicted EONR RMSE between models, Large field (37.3 ha), values are the difference in RMSE (row vs. index) (kg/ha), p-values in parenthese")
```

```{r pair-table-profit, echo=FALSE, cache=FALSE}
model_list <- unique(est_data$model) %>% .[order(.)] %>% as.character()
pair_df <- data.frame()
for(i in 1:length(model_list)){
    for(j in 1:length(model_list)){
        pi <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[i]) %>% 
            pull(profit)
        pj <- est_data %>% 
            filter(field_size=="37.3 ha", model==model_list[j]) %>% 
            pull(profit)
        pv <- t.test(pi, 
                     pj, 
                     paired = TRUE)

        pair_df[model_list[i], model_list[j]] <- 
            round(pv$estimate, 2)
        pair_df[as.character(i), model_list[j]] <- 
            paste0("(", round(pv$p.value, 3), ")")
    }
}
pair_df[is.na(pair_df)] <- ""
knitr::kable(pair_df, format = "markdown",
             caption = "Pairwise t-test of profits between models, Large field (37.3 ha), values are the difference in profits (row vs. index) (kg/ha), p-values in parenthese")
```
